# Understanding time series cross-validation

# Load libraries ----
```{r setup, include=FALSE}
library(dplyr)
library(lubridate)
library(here)
library(timetk)
library(tidymodels)
```

# Create a simple data set with all dates between 2022-01-01 and 2022-12-31
## Assume that the event happened on 1st Jan 2023. And the data from 01-01-2023 is a true hold out.
```{r}
set.seed(0112358)
date_range <- seq(as.Date("2022-01-01"), as.Date("2022-12-31"), by = "day")
dummy_data_simple <- data.frame(date = date_range, hosp_resp = sample(250:525, length(date_range), replace = TRUE))
head(dummy_data_simple)
```

# Display data summary
```{r}
cat("Total observations:", nrow(dummy_data_simple), "\n")
cat("Date range:", as.character(min(dummy_data_simple$date)), "to", as.character(max(dummy_data_simple$date)), "\n")
cat("Hospital respiratory range:", min(dummy_data_simple$hosp_resp), "to", max(dummy_data_simple$hosp_resp), "\n\n")
```

# Step-0: Split data into training and testing
```{r, include=FALSE}
splits_demo <- dummy_data_simple |>
  time_series_split(
    assess = "31 days",
    cumulative = TRUE, # Always true, when using a single slice
    date_var = date
  )
head(training(splits_demo))
# df_training <- training(splits_demo)
# df_testing <- testing(splits_demo)
range(training(splits_demo)$date)
range(testing(splits_demo)$date)
```

# Understanding the parameters of the time_series_cv function using multiple scenarios
## assess: the length of the assessment (validation) period
## initial: the initial training period
## slice_limit: the maximum number of slices to create
## cumulative: whether to use an expanding window

## Function to get the dates of the training and validation periods for each fold
```{r, include=FALSE}
# Function to get CV dates
get_cv_dates <- function(resampling_object) {
  if (!inherits(resampling_object, "time_series_cv")) {
    stop("Input must be a time_series_cv object")
  }
  
  fold_info <- resampling_object$splits |>
    map_dfr(function(split) {
      train_data <- analysis(split)
      val_data <- assessment(split)
      
      train_dates <- train_data |>
        summarise(
          train_start = min(date),
          train_end = max(date),
          train_days = n()
        )
      
      val_dates <- val_data |>
        summarise(
          val_start = min(date),
          val_end = max(date),
          val_days = n()
        )
      
      # Create fold ID - use the split's id if available, otherwise create one
      fold_id <- if (!is.null(split$id)) split$id else "Unknown"
      
      bind_cols(
        id = fold_id,
        train_dates,
        val_dates
      )
    })
  
  fold_info |>
    mutate(
      training_period = paste(format(train_start, "%d-%b-%y"), "to", format(train_end, "%d-%b-%y")),
      validation_period = paste(format(val_start, "%d-%b-%y"), "to", format(val_end, "%d-%b-%y"))
    ) |>
    select(id, training_period, validation_period, train_days, val_days)
}
```

## Scenario 1: Cumulative = FALSE
```{r, include=FALSE}
resamples_demo_s1 <- training(splits_demo) |>
  time_series_cv(
    slice_limit = 5,
    assess = "30 days", # testing period
    skip = "14 days",
    initial = "90 days",
    cumulative = FALSE
  )
```

```{r}
cv_dates_s1 <- get_cv_dates(resamples_demo_s1)
cv_dates_s1
```

- We see that when cumulative = FALSE, the training period is fixed at 30 days (initial)
- The 'initial' parameter is linked to cumulative = FALSE. It is ignored when cumulative = TRUE.
- Also note how the validation period is 7 days for each fold, and they differ by 14 days (skip period)
- Cumulative = FALSE is not useful for time series data, as the training period remians fixed.
- Let us visualize the CV plan

```{r}
plot_s1 <- plot_time_series_cv_plan(resamples_demo_s1, date, hosp_resp)
plot_s1
```

As we can see, a fixed training period trains on different parts of the time series.
This is sub-optimal as it is missing the information from the previous training periods.

## Scenario 2: Cumulative = TRUE
```{r, include=FALSE}
resamples_demo_s2 <- training(splits_demo) |>
  time_series_cv(
    slice_limit = 5,
    assess = "30 days",
    skip = "14 days",
    # initial = "90 days", # This is ignored when cumulative = TRUE
    cumulative = TRUE
  )
```

```{r}
cv_dates_s2 <- get_cv_dates(resamples_demo_s2)
cv_dates_s2
```

Notice how the training period is expanding over the folds, and the validation period is 30 days for each fold and they differ by 14 days (skip period)
Let us visualize the CV plan
```{r}
plot_s2 <- plot_time_series_cv_plan(resamples_demo_s2, date, hosp_resp)
plot_s2
```

The training period is expanding over the folds, while using the entire data available. 
The validation period is 30 days for each fold and they differ by 14 days (skip period)
We can experiment a bit more with the parameters to see how they affect the CV plan.

```{r, include=FALSE}
resamples_demo_s3 <- training(splits_demo) |>
  time_series_cv(
    slice_limit = 20,
    assess = "60 days",
    skip = "30 days",
    # initial = "90 days",
    cumulative = TRUE
  )
```

```{r}
cv_dates_s3 <- get_cv_dates(resamples_demo_s3)
cv_dates_s3
```
